---
title: "Useful Business functions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,roxygen2,scales,glue,dtplyr)
theme_set(theme(axis.text = element_text(color="black")))
install.packages("reprex")
library(reprex)

```




```{r}
#example data
contoso <- readxl::read_excel("Contoso.xlsx",col_names = TRUE)
contoso <- contoso %>% select(-c(contains("DateKey")))
contoso <- janitor::clean_names(contoso)
contoso <- contoso %>% 
  mutate(gross_sale_amount=quantity*unit_price,
           net_sale_amount=quantity*net_price,
           cost_amount=quantity*unit_cost,
           margin=net_sale_amount-cost_amount)
```




```{r}
dput(diamonds)
  
install.packages("SciencesPo")

library(SciencesPo)
```




```{r}
#` creates a customer segmentation
#`
#` @param df is a dataframe or tibble
#` @param group is the grouping key that you want to aggregate the data by, eg customer, plant, product
#` @param dim is the dimension you want to measure, eg. quantity, margin, etc. This should not be a ratio and should be a positive number
#` @param a is the initial threshold for aggregation and should be expressed as a percentage
#` @param b is the incremental threshold for aggregation and should be expressed as a percentage
#` @param c is the incremental threshold for aggregation and should be expressed as a percentage
#` @export

customer_segmentation <- function(df,group,dim,a=.7,b=.26,c=.04){

  
  
  
  
  df %>% #dataframe
    group_by({{group}}) %>% 

    #creates a bunch of columns
    summarize(  
      across({{ dim }}, #make sure this dimension can be aggregratad, later version will handle ratios
             list(sum=~sum(.,na.rm=TRUE),
                  mean=~mean(.,na.rm=TRUE),
                  n =  ~ n(),
                  median =  ~ median(.,na.rm=TRUE),
                  sd =  ~ sd(.,na.rm=TRUE),
                  mad =  ~ mad(.,na.rm=TRUE), #median absolute deviation
                  aad =  ~ mad(., center =mean(.,na.rm=TRUE),na.rm=TRUE), #average absolute deviation
                  IQR05 = ~quantile(., .05,na.rm=TRUE),
                  IQR25 = ~quantile(., .25,na.rm=TRUE),
                  IQR75 = ~quantile(., .75,na.rm=TRUE),
                  IQR95 = ~quantile(., .95,na.rm=TRUE)
                  ),
             .names = "{.fn}") #gives each column their name
    ) %>%
    ungroup() %>% 

    arrange(desc(sum)) %>% # assuming positive values, descends highest to lowest (should add some logic to switch this)
    
    mutate(cum_sum=cumsum(sum), #cumlative value, if ratio, need some sort of check - need specify
         prop_total=sum/max(cum_sum), #assumes positive values, need check
         cum_prop_total=cumsum(prop_total), #cumsum percent of total
         cum_unit_prop=row_number()/max(row_number()), #unit percent
         group_classification_by_dim=
           case_when(
           cum_prop_total<=a ~"A",
           cum_prop_total<=(a+b) ~"B",
           TRUE ~ "C"),
         dim_threshold=
           case_when(group_classification_by_dim=="A"~a,
                     group_classification_by_dim=="B"~(a+b),
                     TRUE ~ c)
         ) %>% 
    select(-c(prop_total,cum_sum)) %>% 
    relocate(dim_threshold,group_classification_by_dim,cum_prop_total,cum_unit_prop)

}

reprex()
```
```{r}


#' Title
#'
#' @param df 
#' @param group 
#' @param dim 
#' @param a 
#' @param b 
#' @param c 
#'
#' @return
#' @export
#'
#' @examples
customer_segmentation <- function(df,group,dim,a=.7,b=.26,c=.04)
  
  
  
  {
  
  if(a+b+c!=1){
    
   stop("a+b+c must equal 1")
    
  } else {
  
  
  
  
  df %>% #dataframe
    group_by({{group}}) %>% 

    #creates a bunch of columns
    summarize(  
      across({{ dim }}, #make sure this dimension can be aggregratad, later version will handle ratios
             list(sum=~sum(.,na.rm=TRUE),
                  mean=~mean(.,na.rm=TRUE),
                  n =  ~ n(),
                  median =  ~ median(.,na.rm=TRUE),
                  sd =  ~ sd(.,na.rm=TRUE),
                  mad =  ~ mad(.,na.rm=TRUE), #median absolute deviation
                  aad =  ~ mad(., center =mean(.,na.rm=TRUE),na.rm=TRUE), #average absolute deviation
                  IQR05 = ~quantile(., .05,na.rm=TRUE),
                  IQR25 = ~quantile(., .25,na.rm=TRUE),
                  IQR75 = ~quantile(., .75,na.rm=TRUE),
                  IQR95 = ~quantile(., .95,na.rm=TRUE)
                  ),
             .names = "{.fn}") #gives each column their name
    ) %>%
    ungroup() %>% 

    arrange(desc(sum)) %>% # assuming positive values, descends highest to lowest (should add some logic to switch this)
    
    mutate(cum_sum=cumsum(sum), #cumlative value, if ratio, need some sort of check - need specify
         prop_total=sum/max(cum_sum), #assumes positive values, need check
         cum_prop_total=cumsum(prop_total), #cumsum percent of total
         cum_unit_prop=row_number()/max(row_number()), #unit percent
         group_classification_by_dim=
           case_when(
           cum_prop_total<=a ~"A",
           cum_prop_total<=(a+b) ~"B",
           TRUE ~ "C"),
         dim_threshold=
           case_when(group_classification_by_dim=="A"~a,
                     group_classification_by_dim=="B"~(a+b),
                     TRUE ~ c)
         ) %>% 
    select(-c(prop_total,cum_sum)) %>% 
    relocate(dim_threshold,group_classification_by_dim,cum_prop_total,cum_unit_prop)

}


```

```{r count_plus}

#count_plus function



count_plus <- function(df,dim,sort=TRUE,...){


df %>% count({{dim}},sort=sort,...) %>% 
  mutate(prop_total=round(cumsum(n/max(cumsum(n))),digits=3),
         unit_prop=round(row_number()/max(row_number()),digits=3),
         row_number=row_number()
         )

}


contoso %>% count_plus(customer_key,wt=margin)

```








```{r}

abc_coord <- function(df){

#create coordinates table----------------------------------------------------------------
df %>% 
  group_by(group_classification_by_dim) %>% 
  summarize(dim_classification_threshold=
              as.numeric(
                round(
                  last(cum_prop_total),
                  digits = 2)
                ),
            cum_by_threshold=
              round(
                last(cum_unit_prop),
                digits = 2),
            n_by=n(),
            plot_cum_dim_prop_total=
              round(first(cum_prop_total),
                    digits = 2),
            plot_cum_by_prop=
              round(first(cum_unit_prop),
                    digits=2),
            ) %>% 
  ungroup() %>% 
    mutate(incr_by_threshold=
             round(
      if_else(
        is.na(
          lag(cum_by_threshold,n = 1)),
        cum_by_threshold,
        cum_by_threshold-lag(cum_by_threshold,n = 1)
        ),
      digits=2))%>% 
    relocate(group_classification_by_dim,
             dim_classification_threshold,
             cum_by_threshold,
             incr_by_threshold,
             n_by)
}

```



```{r}
#create chart-------

abc_graph <- function(df,group_label,dim_label,size,shape,...){

abc_coordinates <- df %>% 
  group_by(group_classification_by_dim) %>% 
  summarize(first_cum_per_of_total=first(cum_prop_total),
            last_cum_per_of_total=last(cum_prop_total),
            first_cum_unit_percent=first(cum_unit_prop),
            last_cum_unit_percent=last(cum_unit_prop),
            n=n(),
            threshold=first(dim_threshold)
            )


 df %>% 
  ggplot(aes(x=cum_unit_prop,
             y=cum_prop_total))+
  geom_point()+
  geom_line()+
  annotate(geom = "rect",
           xmin = abc_coordinates$first_cum_unit_percent,
           xmax = abc_coordinates$last_cum_unit_percent,
           ymin = abc_coordinates$first_cum_per_of_total,
           ymax= abc_coordinates$last_cum_per_of_total,
           alpha=.3,
           fill=c("#007e2f","#ffcd12","#a40000")
           )+
  
#A dotted line---------------------------------------------------------------------------
  
  geom_vline(xintercept = abc_coordinates$last_cum_unit_percent[1], 
             linetype="dashed",
             col="#007e2f")+
#B dotted line--------------------------------------------------------------------------
  
  geom_vline(xintercept = abc_coordinates$last_cum_unit_percent[2],
             linetype="dashed",
             col="#ffcd12")+
  
# C dotted line, I think i take this out???------------------------------------------------------------------------

  geom_vline(xintercept = abc_coordinates$last_cum_unit_percent[3], 
             linetype="dashed",
             col="#a40000")+

# A text -----------------------------------------------------------------------------

  annotate(geom="text", 
           label=
             glue::glue(
               "initial {scales::percent(abc_coordinates$last_cum_unit_percent[1])} ({scales::comma(abc_coordinates$n[1])}) of\n{group_label} drive\n{scales::percent(abc_coordinates$threshold[1])} of {dim_label}"
               ),
           x = abc_coordinates$last_cum_unit_percent[1]+.05,
           y = abc_coordinates$last_cum_unit_percent[1]+.05,
           hjust=0)+
  
# B text ----------------------------------------------------------------------------
  
  annotate(geom="text",
             label=
             glue::glue(
               "cumlative {scales::percent(abc_coordinates$last_cum_unit_percent[2])} of\n{group_label} drive\n{scales::percent(abc_coordinates$threshold[2])} of {dim_label}"
               ),
           hjust=0,
           x =abc_coordinates$last_cum_unit_percent[2]+.05,
           y=abc_coordinates$last_cum_unit_percent[2])+

# C text ----------------------------------------------------------------------

  annotate(geom="text", 
             label=
             glue::glue("remaining {scales::percent(abc_coordinates$threshold[3])} of\n{dim_label} driven by {scales::percent(1-(abc_coordinates$last_cum_unit_percent[1]+abc_coordinates$last_cum_unit_percent[2]))}\n({scales::comma(abc_coordinates$n[3])}) of {group_label}"),
           hjust=0,
           x =abc_coordinates$last_cum_unit_percent[3]-.3,
           y=abc_coordinates$last_cum_unit_percent[3]-.25)+
#format scales---------------------------------------------------------------------------------------------
  scale_x_continuous(labels = scales::percent_format())+
  scale_y_continuous(labels = scales::percent_format())
}

```
\

#distrubition chart
function that takes positive /negative values and makes them both position so that it can be graphed
means, median values of group 1 and overall values 


#box plot
simple box plot (perhaps facet_grid) by group 2

#seasonality

count of transactions in line chart by group 1

##monthly seasonlity 
fill chart by month

## day of month seasonality
fill chart by day of month


## variation table

#group 1 by group 2 or 3

fill chart


#new vs. returning customers

```{r new-vs-returning}
library(glue)

new_vs_returning <- function(df,group,date,new_label,returning_label,date_unit){
  
  df %>%
    group_by({{group}}) %>%
    mutate(first_date=min({{ date}})) %>%
    ungroup() %>% 
    mutate(group_status=
             case_when(
               {{date}}> first_date ~ {{returning_label}},
               {{date}}==first_date ~ {{new_label}},
               TRUE ~ "other")
           ) %>% 
        
    group_by(date_group=
               lubridate::ceiling_date({{date}},unit={{date_unit}})
             ) %>%
    summarize(new=
                n_distinct({{group}}[group_status=={{new_label}}]),
              returning=
                n_distinct({{group}}[group_status=={{returning_label}}]),
            other=
              n_distinct({{group}}[group_status=="other"])
            ) %>% 
  pivot_longer(cols = -c(date_group),names_to = "group_status")
            
    }
  

```


#horizon ridge chart

```{r ggrdiges-chart}

test <- c("hello","ADFDA","123D")


str_view(test,pattern="[1-9]*D")




install.packages("ggridges")

library(ggridges)

contoso %>% count(margin_bucket,sort=T)

contoso$margin_bucket <-cut_number(contoso$margin,6)

contoso %>% 
  ggplot()+
  aes(x=horizon,y=margin_bucket,fill=..x..)+
  ggridges::geom_density_ridges_gradient()+
  MetBrewer::scale_fill_met_c("Cassatt2")

MetBrewer::display_all()

contoso$horizon <- as.numeric(contoso$due_date-contoso$order_date)




```





























graph+
  labs(title="overide")+
  aes(size=n)






  annotate(geom="text", # A text ---------
           label=scales::percent(abc_coordinates$last_cum_unit_percent[1]),
           x =abc_coordinates$last_cum_unit_percent[1]+.05,
           y=A-.1)+
  annotate(geom="text", # B text ----------
           label=scales::percent(abc_coordinates$last_cum_unit_percent[2]),
           x =abc_coordinates$last_cum_unit_percent[2]+.05,
           y=A-.1)+
  annotate(geom="text", # C text -------
           label=scales::percent(1-(abc_coordinates$last_cum_unit_percent[1]+abc_coordinates$last_cum_unit_percent[2])),
           x =abc_coordinates$last_cum_unit_percent[3]+.05,
           y=A-.1)+







customer_ABC_coordinates <- customer_ABC %>% 
  group_by(customer_classification) %>% 
  summarize(first_cum_per_of_total=first(cum_per_of_total),
            last_cum_per_of_total=last(cum_per_of_total),
            first_cum_unit_percent=first(cum_unit_percent),
            last_cum_unit_percent=last(cum_unit_percent)
            )

customer_ABC %>% 
  ggplot(aes(x=cum_unit_percent,y=cum_per_of_total))+
  geom_point()+
  geom_line()+
  annotate(geom = "rect",
           xmin = customer_ABC_coordinates$first_cum_unit_percent,
           xmax = customer_ABC_coordinates$last_cum_unit_percent,
           ymin = customer_ABC_coordinates$first_cum_per_of_total,
           ymax= customer_ABC_coordinates$last_cum_per_of_total,
           alpha=.3,
           fill=c("dark green","yellow","red")
           )+
  geom_vline(xintercept = customer_ABC_coordinates$last_cum_unit_percent[1],
             linetype="dashed",
             col="dark green")+
  annotate(geom="text",
           label=scales::percent(customer_ABC_coordinates$last_cum_unit_percent[1]),
           x =customer_ABC_coordinates$last_cum_unit_percent[1]+.05,
           y=.7-.1)+
  scale_x_continuous(labels = percent_format())+
  scale_y_continuous(labels = percent_format())+
  labs(x="cumlative percent of customers",
       y="",
       title=glue::glue("Cumlaitive margin percent by",{{"product_key"}},"customer rank")
       )+
  theme_classic()

customer_ABC_coordinates
```






*how to checks to ensure total add up to one
*how to ensure packages are there (require)
*how to assign tables or graph to a list



#build custom function


```{r build-function}

ABC <- function(dataset,by,group_by_label,dimension,dimension_label,A=.7,B=.2,C=.1,...){
  
  
  #logic check to ensure values add up 1-----------
  if((A+B+C)!=1){
    stop("categories percentages must sum to 1"
    )
  }
  else{
    #group by dimension and create running totals---------
  dataset_new <- dataset %>% 
    group_by({{ by }}) %>% 
    summarize(sum_dim=sum({{dimension}},na.rm=T),
              mean_dim=mean({{dimension}},na.rm=T),
              n=n()
              ) %>% 
    ungroup() %>% 
    mutate(scale_mean_dim=scale(mean_dim,center = T,scale = T),
           scale_by_dim=scale(n,center = TRUE,scale=TRUE)) %>% 
    arrange(desc(sum_dim)) %>% # assuming positive values, descends highest to lowest
    mutate(cum_sum=cumsum(sum_dim), #cumlative value, if ratio, need some sort of check
         per_of_total=sum_dim/max(cum_sum), #assumes positive values, need check
         cum_per_of_total=cumsum(per_of_total), #cumsum percent of total
         cum_unit_percent=row_number()/max(row_number()), #unit percent
         customer_classification=
           case_when(
           cum_per_of_total<=A ~"A",
           cum_per_of_total<=(A+B) ~"B",
           TRUE ~ "C")
    )
  }
  

#declare coordinates----------
ABC_coordinates <-  dataset_new %>% 
  group_by(customer_classification) %>% 
  summarize(first_cum_per_of_total=first(cum_per_of_total),
            last_cum_per_of_total=last(cum_per_of_total),
            first_cum_unit_percent=first(cum_unit_percent),
            last_cum_unit_percent=last(cum_unit_percent)
            )

#create labels--------


#dimension_label

if (missing(dimension_label)==TRUE){
dimension_label <- dataset %>% select({{dimension}}) %>% colnames() %>% as.character()
  
}
else{
dimension_label <- enquo(dimension_label)
}

#group_by_label

if (missing(group_by_label)==TRUE){
  
group_by_label <- dataset %>% select({{by}}) %>% colnames() %>% as.character()

}

else{

  group_by_label <- enquo(group_by_label)
}


#create chart-------
 dataset_new %>% 
  ggplot(aes(x=cum_unit_percent,y=cum_per_of_total))+
  geom_point()+
  geom_line()+
  annotate(geom = "rect",
           xmin = ABC_coordinates$first_cum_unit_percent,
           xmax = ABC_coordinates$last_cum_unit_percent,
           ymin = ABC_coordinates$first_cum_per_of_total,
           ymax= ABC_coordinates$last_cum_per_of_total,
           alpha=.3,
           fill=c("dark green","yellow","red")
           )+
  geom_vline(xintercept = ABC_coordinates$last_cum_unit_percent[1], #A dotted line---
             linetype="dashed",
             col="dark green")+
  geom_vline(xintercept = ABC_coordinates$last_cum_unit_percent[2], #B dotted line----
             linetype="dashed",
             col="yellow")+
  geom_vline(xintercept = ABC_coordinates$last_cum_unit_percent[3], # C dotted line----
             linetype="dashed",
             col="red")+
  annotate(geom="text", # A text ---------
           label=scales::percent(ABC_coordinates$last_cum_unit_percent[1]),
           x =ABC_coordinates$last_cum_unit_percent[1]+.05,
           y=A-.1)+
  annotate(geom="text", # B text ----------
           label=scales::percent(ABC_coordinates$last_cum_unit_percent[2]),
           x =ABC_coordinates$last_cum_unit_percent[2]+.05,
           y=A-.1)+
  annotate(geom="text", # C text -------
           label=scales::percent(1-(ABC_coordinates$last_cum_unit_percent[1]+ABC_coordinates$last_cum_unit_percent[2])),
           x =ABC_coordinates$last_cum_unit_percent[3]+.05,
           y=A-.1)+
  scale_x_continuous(labels = percent_format())+
  scale_y_continuous(labels = percent_format())+
  labs(x=glue::glue("cumlative percent of ",quo_name(group_by_label)),
       y="",
       title=glue::glue(quo_name(group_by_label)," rank by cumlative ",quo_name(dimension_label))
       )+
  theme_classic()
}

ABC(dataset = contoso,by=customer_key,group_by_label = customer,dimension = margin,A=.7,B=.25,C=.05)

.7+.25+.05


```

```{r count_plus}


count_plus <- function(.data,.var,.dim){

#summaries-------
.data %>% group_by({{.var}}) %>% 
  summarize(n=n(),
            across({{.dim}},median,.names="{.fn}_{.col}"),
            median_var=median(quantity,na.rm = TRUE ),
            mean_var=mean(quantity,na.rm = TRUE),
            sd_var=sd(quantity,na.rm = TRUE),
            mad_var=mad(quantity,
                        center=median(quantity),
                        na.rm = TRUE),
            lad_var=mad(quantity,
                        center=mean(quantity,na.rm = TRUE)),
            min_var=min(quantity,na.rm = TRUE),
            max_var=max(quantity,na.rm = TRUE),
            per_75_var=quantile(quantity,.75,na.rm = TRUE),
            per_25_var=quantile(quantity,.75,),
            mode_var=mode(quantity)
  )
}
?across()


count_plus(.data=contoso,.var=customer_key,.dim=quantity)


#can apply list(.fn1,.fn2,.fn3)


my_summarise <- function(data, group_var) {
  data %>%
    group_by({{ group_var }}) %>%
    summarise(mean = mean(mass))
}
starwars %>% names()

starwars %>% my_summarise(name)

contoso %>% colnames()

count



a <- c(28,30,31)
sum(a)/3
mean(a)
```




1. decide if you have discrete or continuious
2. decide if you want consistency or theme or indifferent

create matrix

1. continous
2. discrete
less than 5
more than 5
more than 10
prgamtically


viridis
rcolorbrewer
hue_pal
grey sclae_


show color: show_col()

color scheme that are gradient and unlimited in number and can be cusotmized
hue_pal

color schemes that are predefined and gradient


color schemes that are discrete and limited in numbers that can be customized


color schemes that are discrete and standalone

how to design your own color scheme

how to rearrange colors?
how to assign colors to factors




time_intelligence <- function(df=datatable,
                              dimension= the dimension to controlfor,
                              metric=the measurement metric,
                              calendar_type=standard or non standard,
                              date_column=which column has the dates associated wth values,
                              period= ,
                              period_match,shift,aggregation,comparison){
  
  
  
  DF %>% 
  date_time() %>% 
  -calendar_type,
  -date_column,
  
  period() %>% 
  shift() %>% 
  aggregation() %>% 
  -metric
  -weights options
  -mark
  -additive
  for totals this will either let you add a total across or will do a different weight
  -granualarity
  For example, if you have monthly values, and someone asks for the monthly average do you take each period and equally weight them? or do you take the daily average balance and multiply it by monthly period average days in each month eg. 28,31,30
  -
  comparison() %>% 
  -compare to what?
  -what to do if unequal measurement lengths



